Fixing Vulnerabilities Potentially Hinders Maintainability

### 2021-02-09 response/response.tex

\end{revcomment}

\Us Thank you for your positive feedback on the paper presentation and replication
package.

We would like to rebut your concern about Better Code Hub (BCH)'s guidelines not 
being related to maintainability. First, as mentioned in the paper, BCH checks GitHub 
codebases against 10 engineering guidelines as devised by Software Improvement Group 
(SIG). These guidelines are the result of many years of experience: analyzing more 
than 15 million lines of code every week, SIG maintains the industry’s largest 
benchmark, containing more than 10 billion lines of code across 200+ technologies; SIG 
is, according to their website, the only lab in the world certified by TÜViT to issue ISO 25010:2011 
certificates for Trusted Product Maintainability\footnote{Information available here: 
\url{https://www.softwareimprovementgroup.com/methodologies/iso-iec-25010-2011-standard/}}.
TÜViT is an IT Security company. Being certified by TÜViT means 
that the company checked if BCH actually respects and follows the ISO 25010 guidelines.

Second, each of the guidelines in isolation is easy to understand, but, when 
combined and evaluated as a whole, it becomes easy to see why 
they would comprise the definition of good software. Furthermore, the relation of the 
guidelines to maintainability is well detailed in the SIG's book ``Building Maintainable 
Software: Ten Guidelines for Future-Proof Code''
\footnote{Available here: \url{https://www.softwareimprovementgroup.com/resources/ebook-building-maintainable-software/}}.
In fact, the underlying maintainability model for BCH is documented with several
(non-academic) reports linked at this page:
\url{https://www.softwareimprovementgroup.com/methodologies/iso-iec-25010-2011-standard/} as 
well as validated by the academic community in the following paper:

@@

\begin{revcomment}[1.4]

    The authors could have adopted both approaches, Better 
    Code Hub guidelines or another one first and then measure 
    the code maintainability with the Delta Maintainability Model 
    since the goal of the paper is to improve security on the 
    maintainability.

\end{revcomment}

\Us As we explained above, namely while addressing comment 1.2, both models assess 
maintainability. We consider the Delta Maintainability 
Model to be an alternative to our model. However, it only uses 5 guidelines of the 
set of guidelines our study considers. This has been 
discussed in the related work section.

@@

\begin{revcomment}[2.1]

    Question:  Is your dataset available?

\end{revcomment}

\Us Thank you for your question. The dataset is fully available at 
\texttt{dataset/db\_security\_changes.csv} in the figshare package
provided in the contributions of the paper: 
\url{https://figshare.com/s/4861207064900dfb3372}. If the paper is 
accepted, the package will be available on GitHub.

This information is described in Section 1 where we present 
the contributions of our work:

\begin{quote}
    This research performs the following main contributions:

\begin{itemize}
  \item Evidence that supports the trade-off between security and 
  maintainability: developers may be hindering software 
  maintainability while patching vulnerabilities.
	\item An empirical study on the impact of security patches on 
	software maintainability (per guideline, severity, weakness and 
	programming language).
	\item A replication package with the scripts and data created to 
	perform the empirical evaluation, for reproducibility. Available 
	online: \url{https://figshare.com/s/4861207064900dfb3372}.
\end{itemize}

\end{quote}

@@

\Us Thank you for raising this issue. In the sentence ``The codebases that resemble to the commits of our datasets 
are collected before the analysis performed by BCH.'', we mean that the codebases of each commit in 
our dataset were analyzed by BCH. We improved the text for:

\begin{quote}
BCH evaluates the codebase available in the default branch of a GitHub project. 
  We created a tool that pulls the codebase of each commit of our dataset 
  to a new branch; it sets the new branch as the default branch; and, runs 
  the BCH analysis on the codebase; after the analysis is finished, the tool saves 
  the BCH metrics results to a cache file.
\end{quote}

\begin{revcomment}[2.9]
    Section 3\\
    \begin{itemize}
        \item Please explain your repeatable methodology to "clean the dataset and select the most relevant and 
        compatible projects" (page 8, line 14) that brings the patches from 1282 to 969.  
        \item From 969 to 866 patches - 
        you just could not classify?  Please explain.
    \end{itemize}
\end{revcomment}

\Us As explained in comment $2.6$, the $1282$ commits refer only to the Ponta et al. dataset (a total of $624$ 
patches). In total, we analyzed $1300$ patches ($624$ patches from Ponta et al. and $676$ from Secbench). 
However, our study only considers $969$ patches. We detected a total of $23$ patches with floss-refactorings. 
In addition, BCH has some limitations, in particular, the lack of language support and project size,
and was uncapable of analyzing the rest of the $308$ data points. All of this is explained
in more detail later in Section 3.4.

@@

\begin{quote}
    A total 
of $103$ patches were not classified because we were not able to map
the issue to any CWE with confidence due to the lack of quality
information on the vulnerability/patch.
\end{quote}

\begin{revcomment}[2.10]
    Section 3\\
    \begin{itemize}
        \item Page 8, line 33 "regular commit" - pick baseline or regular (or non-security) commits and use it always.  
        I don't understand the sentence "The baseline dataset is generated from the security commits dataset."   
        Do you mean you extract these commits from the projects in the security commits dataset?
    \end{itemize}
\end{revcomment}

\Us Yes, for each security commit we collected a random regular change from the same project. We have clarified it 
in the paper.

\begin{quote}
    The baseline dataset is generated from the security commits dataset, i.e.,
for each security commit in the dataset, we collect a random regular
change from the same project. 
\end{quote}

@@

We improved the paper in the following text:

\begin{quote}
    For projects with large codebases, the results retrieved for 
\emph{Keep Your Codebase Small} were not viable because they 
retrieved values above the limit set by BCH ($20$ Person-years). Thus, 
we decided to not consider this guideline in our research. The 
\emph{Automated Tests} guideline was also not considered since the 
tool does not contemplate two of the most important techniques to 
security testing: vulnerability scanning and penetration testing. 
\end{quote}

\begin{revcomment}[2.15]
    Section 3\\
    \begin{itemize}
        \item Page 11  line 48.  How many were disregarded?
    \end{itemize}
\end{revcomment}

\Us Due to BCH limitations, in particular, lack of language support 
and project size by BCH, $308$ data points were disregarded.
We clarified it in the paper (Section 3.4):

\begin{quotation}
    Due to BCH limitations, 
we detected $308$ data points that retrieved incorrect overall 
calculations. Those data points were disregarded from the study.
\end{quotation}

@@

\begin{quote}
    In \emph{RQ2}, we report/discuss the impact of security patches on
    software maintainability per weakness (CWE). We use the weakness definition
    and taxonomy proposed by the \emph{Common Weakness Enumeration} (cf. Section 2).
    Figure 7 shows three different charts. Figure \emph{7-a}, presents
    the impact of the $969$ patches grouped by the first level weaknesses from
    the \emph{Research Concepts} list. While the Figures \emph{7-b} and
    \emph{7-c} present the impact on maintainability for lower levels of 
    weaknesses for the most prevalent weaknesses in Figure \emph{7-a}:
    \emph{Improper Neutralization} (CWE-707) 
    and \emph{Improper Control of a Resource 
    Through its Lifetime} (CWE-664), respectively.

    In Figure~7-\emph{a}, there is no clear evidence of the impact on 
    maintainability per weakness. Yet, it is important to note that
    overall there is a very considerable number of cases that hinder
    maintainability---between $30\%$ and $60\%$. The CWE-707 and CWE-664 
    weaknesses integrate the higher number of cases compared to the remaining 
    ones: $295$ ($30.4\%$) data points and $318$ ($32.8\%$) data points, respectively. 
    Thus, we present an analysis of their sub-weaknesses on 
    Figure~7-\emph{b} and Figure~7-\emph{c}, respectively. 
\end{quote}

\begin{revcomment}[2.19]
    Section 4\\
    \begin{itemize}
        \item Page 20, Line 42 - how many is "a considerable number of cases"?  
    \end{itemize}
\end{revcomment}

\Us In total, we inspected around $25$ regular changes with no impact on software maintainability  
 to understand what was leading to the higher number of cases. This point
 was clarified in Section 4, with the following paragraph:

 \begin{quote}
    Overall, the results for both baselines, show that regular 
    changes are less prone to hinder the software 
    maintainability of open-source software.   
    However, the \emph{size-baseline} integrates a larger 
    number of cases with no impact on software maintainability. 
    We manually inspected a total of $25$ cases from
    that distribution of regular changes with no impact on maintainability, 
    and, found that identifying regular changes with the same 
    size as the security-related commit is limiting the type of regular 
    commits being randomly chosen: input 
    patches, variables or functions, type conversion (i.e., changes
    with no impact on the software metrics analyzed by BCH).
 \end{quote}

@@

\begin{revcomment}[2.21]
    Section 4\\
    \begin{itemize}
        \item RQ3 is answered to succinctly and with too little 
        detail to be convincing to me that the differences between 
        security and regular changes is different.  
    \end{itemize}
\end{revcomment}

\Us We improved the discussion of the comparison between 
the results for security patches and regular changes. We now
consider two baselines: \emph{size-baseline}, where 
we collected random changes with the same size as security 
patches; and, \emph{random-baseline}, where we collected
random changes without any restrictions. The overall results
show that regular changes are less prone to hinder software 
maintainability than security patches. Security
patches results have more cases with negative 
impact on software maintainability than any baseline 
of regular changes. By providing results 
for both baselines, we now clearly see that the limitation 
found for the \emph{size-baseline} does not change 
the need to pay more attention to security patches 
in software maintenance tasks.

\begin{revcomment}[2.22]
    Section 5\\
    \begin{itemize}
        \item Line 32 - since RQ3 was not convincing, 
        I don't feel a separate effort for maintainable security 
        is justified as separate from general focus on maintainability. 
    \end{itemize}
\end{revcomment}

\Us As explained in the previous response, security patches 
hinder more software maintainability than regular changes with size 
restriction (\emph{size-baseline}) or no size restriction (\emph{random-baseline}). Thus, we argue that
special attention should be given to maintainable security
when teaching software maintainability. We improved the analysis 
provided in Section 4 for the research question 3 (RQ3). We added 
an extra baseline for comparison with no size restrictions and 
improved the discussion over our results.

@@

\end{revcomment}

\Us We kindly thank the reviewer for the comments.

As mentioned in the paper, BCH checks GitHub codebases against $10$ engineering 
guidelines as devised by Software Improvement Group (SIG). SIG has devised 
these guidelines after many years of experience: analyzing more than $15$ million 
lines of code every week, SIG maintains the industry’s largest 
benchmark, containing more than $10$ billion lines of code across $200$+ 
technologies; SIG is the only lab in the world certified by TÜViT to issue ISO 25010:2011
certificates for Trusted Product Maintainability\footnote{Information available here: 
\url{https://www.softwareimprovementgroup.com/methodologies/iso-iec-25010-2011-standard/}}.

Each of the guidelines in isolation is easy to understand, but, when 
combined and evaluated as a whole, it becomes easy to see why 
they would comprise the definition of good software. Furthermore, the relation of the entire set of
guidelines to maintainability is well detailed in the SIG's book ``Building Maintainable 
Software: Ten Guidelines for Future-Proof Code''
\footnote{Available here: 
\url{https://www.softwareimprovementgroup.com/resources/ebook-building-maintainable-software/}}.

@@

There are other tools that measure the same metrics such as as Kiuwan; or, even 
provide a model to calculate maintainability such as SonarCloud. In this study, it is not
our goal to study other tools. However, we argue
that BCH is a better tool for the following reasons:

\begin{itemize}
    \item Kiuwan measures 4 guidelines that by name appear to be similar to 
    some of the guidelines measured by BCH: Complexity, Duplicated 
    Code, Size and Coupling\footnote{List available here: \url{https://www.kiuwan.com/docs/display/K5/Metrics}}. 
    However, we did not find the full description of kiuwan's metrics anywhere. Thus, 
    we are ensure that the metrics are the same but we are able to conclude that
    BCH integrates a larger set of metrics.  
    In contrast, 
    SIG provides a full description of the $10$ guidelines/metrics behind better code
    hub plus their relation with maintainability on the book and several scientific papers. 
    \item SonarCloud measures 5 similar metrics to the ones measured by BCH: Complexity, Duplication, 
    Issues, Size and Tests\footnote{List available here: \url{https://sonarcloud.io/documentation/user-guide/metric-definitions/}}. 
    The tool rates maintainability with a letter based on the technical 
    debt ratio which is calculated as the ratio between the cost to develop 
    the software and the cost to fix it. But no information is provided 
    on how the remediation cost and developmet cost are calculated. 
    Better Code Hub contemplates a larger set of maintainability metrics 
    and our formula for maintainability is not only fully explained in our paper but also 
    based on metrics that are fully and publicly described.  
\end{itemize}

We do not know codify and did not find the tool anywhere in static analysis tools 
collections or even through organic search at Google.

We added a Section to 
the paper called \textit{3.3 Better Code Hub} where we motivate the Better Code Hub tool and 
compare it with Kiuwan and SonarCloud.

@@

\begin{revcomment}[3.3]

    The findings of the research should better emphasize what the 
    lessons learned and what better practices should be put in place 
    by developers to assure better quality of software. The take away 
    message remains hindered.

\end{revcomment}

\Us We agree with the reviewer that we should better emphasize the lessons 
learned and what best practices can be followed by developers to assure better 
vulnerability patches. We improved these points in Section 5.

### 2021-02-09 paper/emse20.tex
more fixes:

In this section, we discuss the methodology used to measure
the impact of security patches on the maintainability of open-source
software. The methodology comprises the following steps, as 
illustrated in Figure~\ref{fig:met}.
%
\begin{enumerate}
	\item Combine the datasets from related work that classify
	the activities of developers addressing security-oriented 
  patches~\cite{reis2017secbench,10.1109/MSR.2019.00064}. The
  duplicated patches were tossed (only $2$ were found).
%
	\item
	Extract relevant data (e.g., owner and name of
	the repository, sha key of the vulnerable version, sha key of 	
	the fixed version) from the combined dataset 	
	containing $1300$ security patches collected from open-source 	
	software available on GitHub.
%
  \item Two baselines of regular changes were collected:
  \textit{random-baseline} (for each security commit, a random change 
  was collected from the same project) and \textit{size-baseline} (for 
  each security commit, a random change with the same size
  was collected from the same project). Our goal is to evaluate the impact 
	of regular changes on the maintainability of open-source 
	software.
%
  \item Use the Software Improvement Group (SIG)'s web-based source 
  code analysis service \emph{Better Code Hub} (BCH)
  to quantify maintainability for both security and regular commits. 
  BCH evaluates the codebase available in the default branch of a GitHub project. 
  We created a tool that pulls the codebase of each commit of our dataset 
  to a new branch; it sets the new branch as the default branch; and, runs 
  the BCH analysis on the codebase; after the analysis is finished, the tool saves 
  the BCH metrics results to a cache file.
\end{enumerate}
%
\begin{figure}[h]
	\centering 	
    \begin{adjustwidth}{-0.5cm}{-0.5cm}  
      \vspace{-2.5cm} 
	\includegraphics[width=1.1\textwidth]{figures/methodology.pdf}
  \vspace{-3cm} 
  \caption{Study Methodology}
	\label{fig:met}
	 \end{adjustwidth}
\end{figure}
%
\subsection{Datasets}
%

@@

Overall, the results for both baselines, show that regular 
changes are less prone to hinder the software 
maintainability of open-source software.   
We manually inspected a considerable amount of cases from
the distribution of regular changes that considers size (\emph{size-baseline}), 
where the impact on maintainability 
increases, and, found that identifying regular changes with the same 
size as the security-related commit is limiting the type of regular 
commits being randomly chosen: input 
patches, variables or functions, type conversion, i.e., changes
with no impact on the software metrics analyzed by BCH.
We presume that this phenomenon lead
to the significant number of cases where there is no impact
on the software maintainability. On the other hand, 
identifying regular changes without any restrictions (\emph{random-baseline})
shows that regular changes have less negative impact 
on software maintainability compared with security patches
and that special attention should be given to security patches.

>>>>

In this section, we discuss the methodology used to measure
the impact of security patches on the maintainability of open-source
software. The methodology comprises the following steps, as 
illustrated in Figure~\ref{fig:met}.
%
\begin{enumerate}
	\item Combine the datasets from related work that classify
	the activities of developers addressing security-oriented 
  patches~\cite{reis2017secbench,10.1109/MSR.2019.00064}. The
  duplicated patches were tossed.
%
	\item
	Extract relevant data (e.g., owner and name of
	the repository, sha key of the vulnerable version, sha key of 	
	the fixed version) from the combined dataset 	
	containing $1300$ security patches collected from open-source 	
	software available on GitHub.
%
  \item Two baselines of regular changes were collected:
  \textit{random-baseline} (for each security commit, a random change 
  was collected from the same project) and \textit{size-baseline} (for 
  each security commit, a random change with the same size
  was collected from the same project). Our goal is to evaluate the impact 
	of regular changes on the maintainability of open-source 
	software.
%
  \item Use the Software Improvement Group (SIG)'s web-based source 
  code analysis service \emph{Better Code Hub} (BCH)
  to quantify maintainability for both security and regular commits. 
  BCH evaluates the codebase available in the default branch of a GitHub project. 
  We created a tool that pulls the codebase of each commit of our dataset 
  to a new branch; it sets the new branch as the default branch; and, runs 
  the BCH analysis on the codebase; after the analysis is finished, the tool saves 
  the BCH metrics results to a cache file.
\end{enumerate}
%
\begin{figure}[h]
	\centering 	
    \begin{adjustwidth}{-0.5cm}{-0.5cm}  
      \vspace{-2.5cm} 
	\includegraphics[width=1.1\textwidth]{figures/methodology.pdf}
  \vspace{-3cm} 
  \caption{Study Methodology}
	\label{fig:met}
	 \end{adjustwidth}
\end{figure}
%
\subsection{Datasets}
%

@@

Overall, the results for both baselines, show that regular 
changes are less prone to hinder the software 
maintainability of open-source software.   
However, the \emph{size-baseline} integrates a larger 
number of cases with no impact on software maintainability. 
We manually inspected a total of $25$ cases from
that distribution of regular changes with no impact on maintainability, 
and, found that identifying regular changes with the same 
size as the security-related commit is limiting the type of regular 
commits being randomly chosen: input 
patches, variables or functions, type conversion (i.e., changes
with no impact on the software metrics analyzed by BCH).
We presume that this phenomenon lead
to the significant number of cases where there is no impact
on the software maintainability. On the other hand, 
identifying regular changes without any restrictions (\emph{random-baseline})
shows that regular changes have less negative impact 
on software maintainability compared with security patches
and that special attention should be given to security patches.
